All watched over by machines of loving grace

### 2021-06-22 paper/paper.tex
paper: more work on introduction

Modern interactive theorem proving software---such as Isabelle, Coq, and HOL Light, amongst others---implement a small proof-checking \emph{kernel}.
This kernel is the sole system component that is able to authenticate a purported proof as legitimate, and is typically small and self-contained, making it suitable for manual auditing.
By design, the kernel is the only component within the system that need be trusted by a skeptic, as ultimately all proofs constructed by the system, no matter how large or complex, are eventually passed through the kernel for authentication.
This implementation strategy---commonly known as the LCF-style, after Milner's LCF system that first implemented it---is a reliability mechanism, aiming to ensure that any theorem produced by the system is indeed a theorem.
That is, LCF is a design pattern that helps to write theorem proving software that remains \emph{sound}.

Shifting focus, we see an interesting series of analogies.
Typical commodity operating systems---such as Microsoft Windows and desktop Linux distributions---also implement a small \emph{kernel}.
This kernel is the sole component on the compute system that has unfettered access to hardware.

>>>>

Interactive theorem proving software typically implements a proof-checking \emph{kernel}.
This kernel is the sole system component that is able to authenticate a purported proof as legitimate, and is typically kept small and self-contained, making it suitable for manual auditing.
By design, the kernel is the only component within the system that need be trusted by a skeptic, as all proofs constructed by the system, no matter how large or complex, are ultimately passed through the kernel for authentication.
This implementation strategy---known as the ``LCF-style'', after Milner's system that first implemented it---is a reliability mechanism, aiming to ensure that any theorem produced by the system is indeed a theorem.
In short, the kernel acts as the \emph{root of all trust} within the system.

Shifting focus, we see an interesting analogy.
Typical commodity operating systems---such as Microsoft Windows and desktop Linux distributions---also implement a \emph{kernel}.
This kernel is also the sole system component that has unfettered access to the host machine's hardware\footnote{Barring other privileged systems software, such as the hypervisor, which we consider as another form of operating system}, and will typically provide various abstractions, such as a hierarchically-structured filesystem, sockets, and so on, to untrusted, user-space code.
System-wide security policies are enforced by the kernel using a variety of mechanisms, including password-protected user accounts, file and directory permissions, access control lists, firewalls, and runtime-monitoring facilities.
From a security point-of-view, an operating system kernel again acts as the \emph{root of all trust} within the system.

At heart---though both types of kernels are intended for very different use-cases---both types of kernel are intended to enforce a series of \emph{kernel invariants} even in the face of unbounded interaction with untrusted code.
However, despite this common cause, the implementation techniques employed by authors of the two type of kernels tends to be very different.
In particular, an LCF-style kernel is little other than a library implemented in a high-level programming language---typically some variant of ML, such as Standard ML or OCaml.
Invariants are maintained by the kernel judiciously exposing functionality as an API that offers functions to manipulate elements of a collection of abstract types, thereby exploiting the soundness of the implementation language's type-system.
Untrusted code then interacts with the kernel through this API.
